---
title: "Urban Growth Modeling in Nashville, Tennessee"
author: "Shengqian Wang and Jack Rummler"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tigris)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(sf)
library(raster)
library(knitr)
library(kableExtra)
library(tidycensus)
library(tigris)
library(FNN)
#library(QuantPsyc) # JE Note: in R 4.1, QuantPsyc package not available.
library(caret)
library(yardstick)
library(pscl)
library(plotROC) 
library(ggrepel)
library(pROC)
library(grid)
library(gridExtra)
library(viridis)
library(igraph)
library(RSocrata)

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  panel.border=element_rect(colour="black"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

palette2 <- c("#41b6c4","#253494")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#ffffcc","#a1dab4","#41b6c4","#2c7fb8","#253494")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
```

We also include several helper functions. `quintilesBreaks` takes a dataframe and a column and outputs the quintiles breaks, helping shorten the below `ggplot` calls.

It takes longer to `ggplot` a polygon fishnet with `geom_sf` than it does to plot `geom_point`. To cut down on plotting time, the `xyC` (for ‘XY Coordinates’) takes a fishnet `sf` and converts it to a dataframe of grid cell centroid coordinates.

`rast` is a function allowing us to quickly plot raster values in `ggplot`.

```{r, warning = FALSE, message = FALSE}
#this function converts a column in to quintiles. It is used for mapping.
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }
```

```{r}
tennessee <- counties("TN")

nashville <- tennessee %>%
  dplyr::filter(NAMELSAD == "Davidson County")%>%
  st_transform('ESRI:103527')
ggplot()+
  geom_sf(data=tennessee, fill='grey', color='black')+
  geom_sf(data=nashville, fill='red', color='black')+
  labs(title="Davidson County, Tennessee")+
  mapTheme
```
```{r}
# create fishnet
Nashville_fishnet <- 
  st_make_grid(nashville, 1000) %>%
  st_sf()

# clip to nashville
Nashville_fishnet <-
  Nashville_fishnet[nashville,]
# plot
ggplot() +
  geom_sf(data=Nashville_fishnet) +
  labs(title="Fishnet, 4000 Foot Resolution") +
  mapTheme
```


```{r}
## install.packages("RSocrata")




##  Metropolitan Transit Authority Bus Routes and remove null
bus_stop <- read.socrata(
  "https://data.nashville.gov/resource/vfe9-k7vc.json",
  app_token = "j9XHS8weUKChc4dFGzE03GdIU",
  email     = "ltscqian@gmail.com",
  password  = "qifx!7UR2m4GG8T"
) %>%
  na.omit(bus_stop, cols = "geocoded_column.latitude")



# convert bus_stop to sf+.
bus_stop_sf <- bus_stop %>%
    st_as_sf(coords = c("geocoded_column.longitude", "geocoded_column.latitude"), crs = 4326) %>%
    st_transform(st_crs(nashville)) %>%
    st_intersection(nashville)

#plot 
ggplot()+
  geom_sf(data=nashville, fill='white', color='black')+
  geom_sf(data=bus_stop_sf,color='purple')+
  labs(title="Davidson County, Tennessee")+
  mapTheme
##  bus_route_sf <- bus_route %>%
##    st_as_sf(wkt = "the_geom.corrdiantes", crs = 4326)

```
```{r}
#NLCD LULC & LUCC

lucc_input <- raster("https://github.com/ObjQIAN/warehouseii/raw/main/nlcd_lucc_clip.tif")
lulc_input <- raster("https://github.com/ObjQIAN/warehouseii/raw/main/nlcd_lulc_clip.tif")
plot(lucc_input)

lucc_clip <- mask(lucc_input, nashville)
lulc_clip <- mask(lulc_input, nashville)

#plot lucc
ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lucc_clip) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  labs(title = "Land Cover Change, 2009-2019") +
  mapTheme +
  theme(legend.direction="horizontal")

#plot lulc
ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lulc_clip) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  labs(title = "Land Cover 2011") +
  mapTheme +
  theme(legend.direction="horizontal")

#reclass matrix 
reclassMatrix <- 
  matrix(c(
    0,2,0,
    2,3,1,
    3,Inf,0),
  ncol=3, byrow=T)

reclassMatrix

#reclassify lucc
lucc <- 
  reclassify(lucc_clip,reclassMatrix)

lucc[lucc < 1] <- NA

names(lucc) <- "lc_change"

ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lucc) %>% na.omit, 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange") + 
  labs(title="Development Land Use Change") +
  mapTheme
```

```{r}
changePoints <-
  rasterToPoints(lucc) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(Nashville_fishnet))

fishnet <- 
  aggregate(changePoints, Nashville_fishnet, sum) %>%
  mutate(lc_change = ifelse(is.na(lc_change),0,1),
         lc_change = as.factor(lc_change))

ggplot() +
  geom_sf(data=nashville) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)$x, y=xyC(fishnet)$y, colour=lc_change)) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name = "") +
  labs(title = "Land Cover Development Change", subtitle = "As fishnet centroids") +
  mapTheme
```

