---
title: "Urban Growth Modeling in Nashville, Tennessee"
author: "Shengqian Wang and Jack Rummler"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tigris)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(sf)
library(raster)
library(knitr)
library(kableExtra)
library(tidycensus)
library(tigris)
library(FNN)
#library(QuantPsyc) # JE Note: in R 4.1, QuantPsyc package not available.
library(caret)
library(yardstick)
library(pscl)
library(plotROC) 
library(ggrepel)
library(pROC)
library(grid)
library(gridExtra)
library(viridis)
library(igraph)
library(RSocrata)
library(paletteer)

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  panel.border=element_rect(colour="black"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

qB <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],2),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]],
                                  c(.01,.2,.4,.6,.8), na.rm=T),
                         digits = 3))
  }
}

palette2 <- c("#41b6c4","#253494")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#ffffcc","#a1dab4","#41b6c4","#2c7fb8","#253494")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
               
paletteMain <- c("#AFD0C9FF", "#91A1BAFF", "#51628EFF","#182044FF","#0E122DFF")

paletteQual <- c("#B39DDB", "#9575CD", "#7986CB", "#4FC3F7", "#4DD0E1", "#4DB6AC", "#81C784", "#AED581", "#DCE775", "#FFF176", "#FFD54F", "#FFB74D", "#FF8A65", "#A1887F", "#90A4AE", "#B0BEC5")
```


# Introduction

Urban growth forecasting is a tactful strategy to predict where people may live in the future. Looking at population projections and land cover data, we can understand areas that may have greater development pressure so planners can prepare for future growth scenarios. In this study, we forecast urban growth in Davidson County, Tennessee.

```{r, fig.width=12}
tennessee <- counties("TN") # county via tigris

nashville <- tennessee %>%
  dplyr::filter(NAMELSAD == "Davidson County")%>%
  st_transform('ESRI:103527')



ggplot()+
  geom_sf(data=tennessee, fill='grey', color='black')+
  geom_sf(data=nashville, fill='red', color='black')+
  labs(title="Davidson County, Tennessee")+
  mapTheme()
```

## Study Region

Davidson County is the second-largest county in the state of Tennessee, with Nashville as the city capital. We chose this our study region for urban growth modeling given that between 2010 and 2020, the county grew over 11%, which is well above the national average. We wanted to select a study region that still had room to develop, but is experiencing extensive population growth, indicating both the need for supply and demand side solutions for future development.

## Methodology

We are developing a predictive model of urban growth for the year 2029. We do this by collecting land use cover change from the years 2011 and 2019 to analyze change between the two, and predict the land use change between 2019 and 2029. We also engineer a variety of features ranging from land use, population, infrastructure, and development. Land cover change, our dependent variable, and our predictive variables will be aggregated into a fishnet of our study region to create a granular, equal spatial structure. We will then use this model to develop an allocation procedure from both the demand side and supply side of future development.

**Demand:** Looking at population projections for the year 2029, we distribute population projections among Davidson County's and predict land cover and future development in 2029.
**Supply:** xxx

## Planning Application

Planners often try to assess population growth using a variety of strategies, such as demographic trend analyses, economic analyses, and scenario planning, to name a few. We are completing this model because the process of predicting growth is challenging. While using previous population data, historical land use change, and other spatially related factors, growth of a city is often non-linear and non-predictive given factors like economic development and job opportunities, public polcicies, and migration. 

Moreover, while regions like Davidson County have a strong urban core, much of the region is rapidly sprawling. In 2014, a [study]("https://wpln.org/post/how-bad-is-nashvilles-sprawl-new-study-ranks-it-among-the-worst/") by Smart Growth America and University of Utah ranked the Nashville region as one of the worst of nearly 221 metro areas for urban sprawl. We know that incentivizing growth in areas with infill potential, investing in transit-oriented development, and disincentivizing continual urban sprawl can help planners engage in more sustainable, compact development to minimize strain of infrastructure and natural resources.

```{r, warning = FALSE, message = FALSE}
#this function converts a column in to quintiles. It is used for mapping.
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }
```

# Data Set-Up

## Fishnet

We first create the spatial structure of our fishnet grid. We decided to use a 1000 meter resolution. This was a challenge to decide given the balance between maximizing our model's accuracy while balancing out the computation intensity of the modeling process.

```{r}
# create fishnet
Nashville_fishnet <- 
  st_make_grid(nashville, 1000) %>%
  st_sf()

# clip to nashville
Nashville_fishnet <-
  Nashville_fishnet[nashville,]

# plot
ggplot() +
  geom_sf(data=Nashville_fishnet) +
  labs(title="Fishnet, 1000 Meter Resolution") +
  mapTheme()
```

```{r}
Nashville_fishnet <-
  Nashville_fishnet %>%
  rownames_to_column("fishnetID") %>% 
  mutate(fishnetID = as.numeric(fishnetID)) %>%
  dplyr::select(fishnetID)
```

## Land Use Change

```{r}
#NLCD LULC & LUCC

lucc_input <- raster("https://github.com/ObjQIAN/warehouseii/raw/main/LUCC_2019_11.tif")
lulc_input <- raster("https://github.com/ObjQIAN/warehouseii/raw/main/nlcd_lulc_clip.tif")
#plot(lucc_input)

lucc_clip <- mask(lucc_input, nashville)
lulc_clip <- mask(lulc_input, nashville)

#plot lucc
ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lucc_clip) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_manual(values=c(paletteQual))+
  labs(title = "Land Cover Change, 2009-2019") +
  mapTheme() +
  theme(legend.direction="horizontal")
```

```{r}
#plot lulc
ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lulc_clip) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  labs(title = "Land Cover (2011)", subtitle="Davidson County, TN") +
  scale_fill_manual(values=c(paletteQual), name="Land Cover Type")+
  mapTheme() +
  theme(legend.position="bottom")
```

```{r}
#reclass matrix 
reclassMatrix <- 
  matrix(c(
    0,12,0,
    12,24,1,
    24,Inf,0),
  ncol=3, byrow=T)

reclassMatrix
```

```{r}
#reclassify lucc
lucc <- 
  reclassify(lucc_clip,reclassMatrix)

lucc[lucc < 1] <- NA

names(lucc) <- "lc_change"

ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lucc) %>% na.omit, 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange") + 
  labs(title="Development Land Use Change") +
  mapTheme()



```

```{r}
changePoints <-
  rasterToPoints(lucc) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(Nashville_fishnet))

fishnet <- 
  aggregate(changePoints, Nashville_fishnet, sum) %>%
  mutate(lc_change = ifelse(is.na(lc_change),0,1),
         lc_change = as.factor(lc_change))

ggplot() +
  geom_sf(data=nashville) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)$x, y=xyC(fishnet)$y, colour=lc_change)) +
  scale_colour_manual(values = c("#4FC3F7", "#81C784"),
                      labels=c("No Change","New Development"),
                      name = "") +
  labs(title = "Land Cover Development Change", subtitle = "As fishnet centroids") +
  mapTheme()
```

```{r}
## plot land use in 2011
ggplot() +
  geom_sf(data=nashville) +
  geom_raster(data=rast(lulc_clip) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="") +
  labs(title = "Land Cover, 2011") +
  mapTheme() +
  theme(legend.direction="horizontal")
```

```{r}
# landuse reclassification
developed <- lulc_clip == 21 | lulc_clip == 22 | lulc_clip == 23 | lulc_clip == 24
forest <- lulc_clip == 41 | lulc_clip == 42 | lulc_clip == 43 
farm <- lulc_clip == 81 | lulc_clip == 82 
wetlands <- lulc_clip == 90 | lulc_clip == 95 
otherUndeveloped <- lulc_clip == 52 | lulc_clip == 71 | lulc_clip == 31 
water <- lulc_clip == 11

names(developed) <- "developed"
names(forest) <- "forest"
names(farm) <- "farm"
names(wetlands) <- "wetlands"
names(otherUndeveloped) <- "otherUndeveloped"
names(water) <- "water"


```

```{r}
# aggregateRaster

aggregateRaster <- function(inputRasterList, theFishnet) {
  #create an empty fishnet with the same dimensions as the input fishnet
  theseFishnets <- theFishnet %>% dplyr::select()
  #for each raster in the raster list
  for (i in inputRasterList) {
  #create a variable name corresponding to the ith raster
  varName <- names(i)
  #convert raster to points as an sf
    thesePoints <-
      rasterToPoints(i) %>%
      as.data.frame() %>%
      st_as_sf(coords = c("x", "y"), crs = st_crs(theFishnet)) %>%
      filter(.[[1]] == 1)
  #aggregate to the fishnet
    thisFishnet <-
      aggregate(thesePoints, theFishnet, length) %>%
      mutate(!!varName := ifelse(is.na(.[[1]]),0,1))
  #add to the larger fishnet
    theseFishnets <- cbind(theseFishnets,thisFishnet)
  }
  #output all aggregates as one large fishnet
   return(theseFishnets)
  }
```

# Feature Aggregation

The following features have been added to our fishnet below, in addition to our land use change:

1. Total Population - American Community Survey (2011 and 2019)
2. % of Occupied Housing Units - American Community Survey (2011 and 2019)
3. Building Permit Applications - Open Data Nashville (2020 to 2023)
...

## Census Data

We pulled two main variables from the census: total population and occupied housing units. We hypothesized that these were predictive of development, thus predictive of land use change. With a greater density of people and housing units, these areas are likely to grow. We gathered these variables at the census tract level.

```{r}
census_api_key("3c896250ea8d421462ade754e4dcecdf8f55e0f2", overwrite = TRUE)

nashvillePop2011 <- 
  get_acs(geography = "tract", 
                variables = c("B01001_001E", # Total population
                              "B25002_002E" # Occupied housing units
                ),
                year = 2011,
                state = 47, 
                geometry = TRUE, 
                county = 037,
                output = "wide") %>%
  rename(totalPop2011 = B01001_001E,
         totalOccupied2011 = B25002_002E
         ) %>%
  dplyr::select(-c("B01001_001M", "B25002_002M")) %>%
  st_transform('ESRI:103527')

nashvillePop2019 <- 
   get_acs(geography = "tract", 
                variables = c("B01001_001E", # Total population
                              "B25002_002E" # Occupied housing units
                ),
                year = 2019,
                state = 47, 
                geometry = TRUE, 
                county = 037,
                output = "wide") %>%
  rename(totalPop2019 = B01001_001E,
         totalOccupied2019 = B25002_002E
         ) %>%
  dplyr::select(-c("B01001_001M", "B25002_002M")) %>%
  st_transform('ESRI:103527')
```

### Total Populaion - 2011 vs. 2019

First, we look at total population within each census tract in years 2011 and 2019.

```{r, fig.width=12}
grid.arrange(
ggplot()+
  geom_sf(data=nashvillePop2011, aes(fill=q5(totalPop2011)), color=NA) +
  scale_fill_manual(values=paletteMain,
                    labels=(qBr(nashvillePop2011, "totalPop2011")),
                    name="Population\nQuintiles")+
  labs(title = "Total Population - 2011",
       subtitle = "Davidson County Census Tracts")+
  mapTheme(),
ggplot()+
  geom_sf(data=nashvillePop2019, aes(fill=q5(totalPop2019)), color=NA) +
  scale_fill_manual(values=paletteMain,
                    labels=(qBr(nashvillePop2019, "totalPop2019")),
                    name="Population\nQuintiles")+
  labs(title = "Total Population - 2019",
       subtitle = "Davidson County Census Tracts")+
  mapTheme(),
ncol=2
)
```

### Occupied Housing Units - 2011 vs. 2019

Now, we visualize the total occupied housing units in each census tract.

```{r, fig.width=12}
grid.arrange(
ggplot()+
  geom_sf(data=nashvillePop2011, aes(fill=q5(totalOccupied2011)), color=NA) +
  scale_fill_manual(values=paletteMain,
                    labels=(qB(nashvillePop2011, "totalOccupied2011")),
                    name="Housing Unit\nQuintiles")+
  labs(title = "Occupied Housing Units - 2011",
       subtitle = "Davidson County Census Tracts")+
  mapTheme(),
ggplot()+
  geom_sf(data=nashvillePop2019, aes(fill=q5(totalOccupied2019)), color=NA) +
  scale_fill_manual(values=paletteMain,
                    labels=(qB(nashvillePop2019, "totalOccupied2019")),
                    name="Housing Unit\nQuintiles")+
  labs(title = "Occupied Housing Units - 2019",
       subtitle = "Davidson County Census Tracts")+
  mapTheme(),
ncol=2
)
```

### Census Features as Fishnet

We aggregate all of the census features for each year into our fishnet grid, and visualize them within the fishnet, also including the total change in population and housing units for each grid below.

```{r, fig.width=12}
fishnetPopulation11 <-
  st_interpolate_aw(nashvillePop2011["totalPop2011"], Nashville_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(Nashville_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(totalPop2011 = replace_na(totalPop2011,0)) %>%
  dplyr::select(totalPop2011)

fishnetPopulation19 <-
  st_interpolate_aw(nashvillePop2019["totalPop2019"], Nashville_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(Nashville_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(totalPop2019 = replace_na(totalPop2019,0)) %>%
  dplyr::select(totalPop2019)

fishnetHousing11 <-
  st_interpolate_aw(nashvillePop2011["totalOccupied2011"], Nashville_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(Nashville_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(totalOccupied2011 = replace_na(totalOccupied2011,0)) %>%
  dplyr::select(totalOccupied2011)

fishnetHousing19 <-
  st_interpolate_aw(nashvillePop2019["totalOccupied2019"], Nashville_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(Nashville_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(totalOccupied2019 = replace_na(totalOccupied2019,0)) %>%
  dplyr::select(totalOccupied2019)

fishnetCensus <- 
  cbind(fishnetPopulation11,fishnetPopulation19, fishnetHousing11, fishnetHousing19) %>%
  dplyr::select(totalPop2011, totalPop2019, totalOccupied2011, totalOccupied2019) %>%
  mutate(popChange = (totalPop2019 - totalPop2011),
         housingChange = (totalOccupied2019 - totalOccupied2011))

grid.arrange(
ggplot()+
  geom_sf(data=fishnetCensus, aes(fill=q5(totalPop2011)), color=NA)+
  scale_fill_manual(values=paletteMain,
                    labels=(qB(fishnetCensus, "totalPop2011")),
                    name="Quintiles")+
  labs(title="Fishnet: 2011 Population")+
  mapTheme(),
ggplot()+
  geom_sf(data=fishnetCensus, aes(fill=q5(totalPop2019)), color=NA)+
  scale_fill_manual(values=paletteMain,
                    labels=(qB(fishnetCensus, "totalPop2019")),
                    name="Quintiles")+
  labs(title="Fishnet: 2019 Population")+
  mapTheme(),
ggplot()+
  geom_sf(data=fishnetCensus, aes(fill=q5(totalOccupied2011)), color=NA)+
  scale_fill_manual(values=paletteMain,
                    labels=(qB(fishnetCensus, "totalOccupied2011")),
                    name="Quintiles")+
  labs(title="Fishnet: 2011 Occupied Housing Units")+
  mapTheme(),
ggplot()+
  geom_sf(data=fishnetCensus, aes(fill=q5(totalOccupied2019)), color=NA)+
  scale_fill_manual(values=paletteMain,
                    labels=(qB(fishnetCensus, "totalOccupied2019")),
                    name="Quintiles")+
  labs(title="Fishnet: 2019 Occupied Housing Units")+
  mapTheme(),
ggplot()+
  geom_sf(data=fishnetCensus, aes(fill=q5(popChange)), color=NA)+
  scale_fill_manual(values=paletteMain,
                    labels=(qB(fishnetCensus, "popChange")),
                    name="Quintiles")+
  labs(title="Fishnet: Population Change")+
  mapTheme(),
ggplot()+
  geom_sf(data=fishnetCensus, aes(fill=q5(housingChange)), color=NA)+
  scale_fill_manual(values=paletteMain,
                    labels=(qB(fishnetCensus, "housingChange")),
                    name="Quintiles")+
  labs(title="Fishnet: Occupied Housing Units Change")+
  mapTheme(),
ncol=2
)
```

# Open Data Nashville

## Building Permit Applications

```{r}

buildingPermits <- read.socrata(
  "https://data.nashville.gov/resource/kqff-rxj8.json",
  app_token = "j9XHS8weUKChc4dFGzE03GdIU",
  email     = "ltscqian@gmail.com",
  password  = "qifx!7UR2m4GG8T"
)

buildingPermits <- buildingPermits[complete.cases(buildingPermits$mapped_location.latitude),]
buildingPermits <- buildingPermits %>% dplyr::filter(address != "115  GREAT CIRCLE RD")
buildingPermits <- st_as_sf(buildingPermits, 
                            coords = c("mapped_location.longitude", "mapped_location.latitude"), crs = 4326)

buildingPermits <- buildingPermits %>% st_transform('ESRI:103527')

ggplot()+
  geom_sf(data=nashville, fill='grey', color='black', size=20)+
  geom_sf(data=buildingPermits, alpha=0.4)+
  labs(title="Building Permit Requests",
       subtitle="Davidson County, TN (2020 to 2023)")+
  mapTheme()
```

```{r}
fishnetPermits <-
  dplyr::select(buildingPermits) %>% 
  mutate(countPermits = 1) %>% 
  aggregate(., Nashville_fishnet, sum) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(Nashville_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  dplyr::select(countPermits)

fishnetPermits$countPermits <- ifelse(is.na(fishnetPermits$countPermits), 0, fishnetPermits$countPermits)

colorRampMain <- colorRampPalette(paletteMain)
colorRampMain5 <- colorRampMain(5)

ggplot()+
  geom_sf(data=fishnetPermits, aes(fill=as.numeric(countPermits)), color=NA)+
  scale_fill_viridis(option="G", direction=-1,
                     name="Count") +
  labs(title="Fishnet: Count of Building Permits")+
  mapTheme()

```


```{r}
theRasterList <- c(developed,forest,farm,wetlands,otherUndeveloped,water)

aggregatedRasters <-
  aggregateRaster(theRasterList, Nashville_fishnet) %>%
  dplyr::select(developed,forest,farm,wetlands,otherUndeveloped,water) %>%
  mutate_if(is.numeric,as.factor)

aggregatedRasters %>%
  gather(var,value,developed:water) %>%
  st_cast("POLYGON") %>%    #just to make sure no weird geometries slipped in
  mutate(X = xyC(.)$x,
         Y = xyC(.)$y) %>%
  ggplot() +
    geom_sf(data=nashville) +
    geom_point(aes(X,Y, colour=as.factor(value))) +
    facet_wrap(~var) +
    scale_colour_manual(values = palette2,
                        labels=c("Other","Land Cover"),
                        name = "") +
    labs(title = "Land Cover Types, 2001",
         subtitle = "As fishnet centroids") +
   mapTheme()
```

```{r}
## install.packages("RSocrata")

##  Metropolitan Transit Authority Bus Routes and remove null
bus_stop <- read.socrata(
  "https://data.nashville.gov/resource/vfe9-k7vc.json",
  app_token = "j9XHS8weUKChc4dFGzE03GdIU",
  email     = "ltscqian@gmail.com",
  password  = "qifx!7UR2m4GG8T"
) %>%
  na.omit(bus_stop, cols = "geocoded_column.latitude")



# convert bus_stop to sf+.
bus_stop_sf <- bus_stop %>%
    st_as_sf(coords = c("geocoded_column.longitude", "geocoded_column.latitude"), crs = 4326) %>%
    st_transform(st_crs(nashville)) %>%
    st_intersection(nashville)

#plot 
ggplot()+
  geom_sf(data=nashville, fill='white', color='black')+
  geom_sf(data=bus_stop_sf,color='purple')+
  labs(title="Davidson County, Tennessee")+
  mapTheme()
##  bus_route_sf <- bus_route %>%
##    st_as_sf(wkt = "the_geom.corrdiantes", crs = 4326)

```

```{r}

# plot stops with fishnet, do not know if it will work(see urban growth rmd)
ggplot() +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2],colour=lc_change),size=1.5) +
  geom_sf(data=bus_stop_sf) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development")) +
  labs(title = "New Development and Bus stops",
       subtitle = "As fishnet centroids") +
  mapTheme()

```

```{r}
# measure distance to nearest bus stop

#emptyRaster <- lucc
#emptyRaster[] <- NA
# number coule be desired factor to reduce resolution
lower_res_raster <- aggregate(lucc, fact = 5) 
lower_res_raster[] <- NA

bus_sf_geo_only <- bus_stop_sf %>% select(geometry)# remove all the columns except for geometry (or will be error in rasterize step)

bus_stop_raster <- 
  as(bus_sf_geo_only,'Spatial') %>%
  rasterize(.,lower_res_raster)

bus_stop_distance <- distance(bus_stop_raster)
names(bus_stop_distance) <- "distance_stops"

busstop_Points <-
  rasterToPoints(bus_stop_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(Nashville_fishnet))

busstop_Points_fishnet <- 
  aggregate(busstop_Points, Nashville_fishnet, mean) %>%
  mutate(distance_stops = ifelse(is.na(distance_stops),0,distance_stops))

ggplot() +
  geom_sf(data=nashville) +
  geom_point(data=busstop_Points_fishnet, aes(x=xyC(busstop_Points_fishnet)[,1], 
                                             y=xyC(busstop_Points_fishnet)[,2], 
                 colour=factor(ntile(distance_stops,5))),size=1.5) +
  scale_colour_manual(values = palette5,
                      labels=substr(quintileBreaks(busstop_Points_fishnet,"distance_stops"),1,8),
                      name="Quintile\nBreaks") +
  geom_sf(data=bus_stop_sf, colour = "red") +
  labs(title = "Distance to Bus stops",
       subtitle = "As fishnet centroids; Bus stops visualized in red") +
  mapTheme()
```


## spatial lag

```{r}
#define function
nn_function <- function(measureFrom,measureTo,k) {
  #convert the sf layers to matrices
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}

fishnet$lagDevelopment <-
    nn_function(xyC(fishnet),
                xyC(filter(aggregatedRasters,developed==1)),
                2)

ggplot() +
  geom_sf(data=nashville) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2], 
                 colour=factor(ntile(lagDevelopment,5))), size=1.5) +
  scale_colour_manual(values = palette5,
                     labels=substr(quintileBreaks(fishnet,"lagDevelopment"),1,7),
                     name="Quintile\nBreaks") +
  labs(title = "Spatial Lag to 2001 Development",
       subtitle = "As fishnet centroids") +
  mapTheme()
``` 
## fixed effect - cropland
```{r}


crop_VALUE1 <- 81
crop_VALUE2 <- 82

# Define a function to keep only crop value, and set other values to 0
keep_values <- function(x) {
  x[!(x %in% c(crop_VALUE1, crop_VALUE2))] <- NA
  return(x)
}

# Apply the function to the original raster
cropland_raster <- calc(lulc_clip, keep_values)

#cropland_raster <- aggregate(lulc_clip, fact = 5) 
cropland_raster <- mask(cropland_raster, nashville)
cropland_raster <- aggregate(cropland_raster, fact = 5) 
cropland_distance <- distance(cropland_raster)
names(cropland_distance) <- "distance_croplands"

cropland_Points <-
  rasterToPoints(cropland_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(Nashville_fishnet))

cropland_Points_fishnet <- 
  aggregate(cropland_Points, Nashville_fishnet, mean) %>%
  mutate(distance_croplands = ifelse(is.na(distance_croplands), 0, distance_croplands))



ggplot() +
  geom_sf(data=nashville) +
  geom_point(data=cropland_Points_fishnet, aes(x=xyC(cropland_Points_fishnet)[,1], 
                                             y=xyC(cropland_Points_fishnet)[,2], 
                 colour=factor(ntile(distance_croplands, 5))), size=1.5) +
  scale_colour_manual(values = palette5,
                      labels=substr(quintileBreaks(cropland_Points_fishnet, "distance_croplands"), 1, 8),
                      name="Quintile\nBreaks") +
#  geom_raster(data=cropland_raster, aes(x, y, fill=as.factor(value)), alpha=0.5) + # Add the cropland raster to the plot
  labs(title = "Distance to Croplands",
       subtitle = "As fishnet centroids; Croplands visualized in red") +
  mapTheme()
```

